---
title: "Reproducible_code"
author: "T Ratnaike"
date: "2023-05-04"
output: html_document
---
```{r, include=FALSE}
download.file(url="https://raw.githubusercontent.com/cambiotraining/reproducibility-training/master/data/gapminder_data.csv", destfile="~/Desktop/Paediatric ST5/Research/data/gapminder_data.csv")
```

```{r load data, echo=FALSE, message= FALSE, warning= FALSE}
library(tidyverse)
library(rmarkdown)
library(kableExtra)
library(ggpubr)

pop_data <- read_csv("~/Desktop/Paediatric ST5/Research/Training/data/gapminder_data.csv")

euro_data_tbl <- pop_data %>% 
  filter(continent== "Europe" & year == 2007) %>%
  select(-continent, -year) %>% 
  arrange(desc(lifeExp)) %>% 
  rename(Country = country, 
         "Population Size" = pop, 
         "Life Expectancy" = lifeExp, 
         "GDP" = gdpPercap)
  

```

```{r kbl}
euro_data_tbl %>% kable(caption="European countries ordered by greatest life expectancy from 2007 data") %>% 
          kable_styling(bootstrap_options = "striped", full_width = F, "condensed") %>% 
          scroll_box(width = "100%", height = "200px")
```

#A better way to display this table is with pagination as follows:

```{r paged-table, echo=FALSE}
paged_table(euro_data_tbl)
```

Next, the life expectancy in Europe is observed across different years.  The aim of this is to check if there was a change in life expectancy over the years.  The data used for the life expectancy over the years analysis is the gapminder data and the following data manipulation was performed:

```{r fig-data}
#keep on European data and change year to factor
euro_data_fig <- pop_data %>% 
                    filter(continent == "Europe")  %>% 
                    mutate(year=as_factor(year))

```

The life expectancy for Europe over the years is plotted as following:

```{r fig, warning=FALSE, fig.cap="Life Expectancy in Europe over the years"}
#Euro plot
euro_plot <- euro_data_fig %>%
                ggplot(mapping=aes(x=year, y=lifeExp)) +
                geom_violin() +
                stat_summary(fun.y = median, geom = "point")

#draw euro plot
euro_plot

```
